Spanner: Google's Globally-Distributed Database \newline Seminar Report

### 2016-02-16 paper/paper.tex
Iterate Paxos Leader Lease
Iterate Read-Write Transactions

\subsection{Paxos Leader Leases \tbr}
\label{subsec:paxos-leader-leases}

The Paxos implementation of Spanner is designed for long-lived leaderships. As
usual in Paxos implementations, a leader is elected by a quorum of lease votes.
If a machine receives such a quorum of lease votes, it gets a lease (10 seconds
by default). The lease vote of a replica is extended automatically on successul
writes.

@@

\subsection{Read-Write Transactions \tbr}
\label{subsec:read-write-transactions}

Read-write transaction use 2-phase locking and Spanner may only assign timestamps
in between the two phases. Spanner enforces another invariant, i.e. a
\emph{monotonicity invariant}, which makes use of the disjointness invariant.
This monotonicity invariant guarantees that within each Paxos group, Spanner
assigns timestamps to Paxos writes in monotonically increasing order. To guarantee
this across leaders, Spanner utilizes the disjointness invariant. This implies
that leaders must only assign timestamps within the interval of their respective
leader leases. $s_{max}$ is advanced to $s$ whenever a timestamp $s$ is assigned
by a leader.

To guarantee externally-consistent reads and writes, Spanner enforces a
\emph{external-consistency invariant}. In essence, this invariant guarantees that
whenever the commit of a transaction $T_1$, denoted as $e_1^{commit}$, occurs
before the start of a transaction $T_2$, denoted as $e_2^{start}$, then the commit
timestamp of $T_1$, denoted as $s_1$, is guaranteed to less than the timestamp
of $T_2$, denoted as $s_2$. More formally, the following must hold to guarantee
external consistency:
\begin{center}
$t_{abs}\left(e_1^{commit}\right) < t_{abs}\left(e_2^{start}\right) \Rightarrow s_1 < s_2$.
\end{center}

Spanner obeys two rules in order to guarantee this invariant:
\begin{description}
  \item[Start] Coordinator leader for a write $T_i$ assigns a commit timestamp
    $s_i \geq TT.now().latest$, where $TT.now()$ is executed right after the
    arrival event of commit request for $T_i$.
  \item[Commit Wait] Coordinator leader ensures that clients cannot see data
    committed by $T_i$ until $TT.after\left(s_i\right)$ is true. This ensures
    that the assigned timestamp $s_i$ is less than the absolute commit time of
    $T_i$.~\cite{Corbett:2012}
\end{description}

Figure~\ref{fig:read-write-transaction} illustrates the timeline of a distributed
read-write transaction. There are three participants groups involved, i.e. $P_2$,
$P_3$ and $P_{CL}$ (which is the coordinator leader). In the first step, all three
participants acquire their locks and execute $TT.now()$ right afterwards to obtain
timestamps independently. Every participant logs his timestamp through Paxos and
after this was done, sends its timestamp to the coordinator leader. The
coordinator leader then computes an overall timestamp for the transaction. This
timestamp must be greater than any timestamp the leader has assigned before,
greater or equal to all timestamps received by the participants and greater than
$TT.now().latest$ at the time the coordinator leader received the commit message.
Then, the computed timestamp is logged by the coordinator leader through Paxos
but no participants are notified until the commit wait rule is satisfied. After
commit wait is satisfied, the coordinator leader notifies all other participants
which can release their locks after receiving the notification.

@@

    \node[right = -0.15 of notify-t2] (release-locks-t2) {};
    \node[right = -0.15 of notify-t3] (release-locks-t3) {};
    \draw[red] (release-locks-t2.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t2.center) -- ++(0, 0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, 0.25);
  \end{tikzpicture}
  \caption{Read-write transaction}
  \label{fig:read-write-transaction}
\end{figure}

This also explains why, according to Table~\ref{tbl:reads-writes}, Read-write
transactions only involve the \emph{leader} replicas. This is because only the
leader replicas of each participant communicate with the coordinator leader.
Hence, no other replicas are involved.

>>>>

\subsection{Paxos Leader Leases}
\label{subsec:paxos-leader-leases}

The Paxos implementation of Spanner is designed for long-lived leaderships. As
usual in Paxos implementations, a leader is elected by a quorum of lease votes.
If a machine receives such a quorum of lease votes, it gets a lease (10 seconds
by default). The lease vote of a replica is extended automatically on successful
writes.

@@

\subsection{Read-Write Transactions}
\label{subsec:read-write-transactions}

Read-write transactions use two-phase locking and Spanner may only assign
timestamps in between the two phases. Spanner enforces another invariant, i.e. a
\emph{monotonicity invariant}, which makes use of the disjointness invariant.
This monotonicity invariant guarantees that within each Paxos group, Spanner
assigns timestamps to Paxos writes in monotonically increasing order. To guarantee
this across leaders, Spanner utilizes the disjointness invariant. This implies
that a leader must only assign timestamps within the interval of its leader
lease. $s_{max}$ is advanced to $s$ whenever a timestamp $s$ is assigned by a
leader.

To guarantee externally-consistent reads and writes, Spanner enforces an
\emph{external consistency invariant}. In essence, this invariant guarantees that
whenever the commit of a transaction $T_1$, denoted as $e_1^{commit}$, occurs
before the start of a transaction $T_2$, denoted as $e_2^{start}$, then the commit
timestamp of $T_1$, denoted as $s_1$, is guaranteed to be smaller than the
timestamp of $T_2$, denoted as $s_2$. More formally, the following must hold to
guarantee external consistency:
\begin{center}
$t_{abs}\left(e_1^{commit}\right) < t_{abs}\left(e_2^{start}\right) \Rightarrow s_1 < s_2$
\qquad\qquad $t_{abs}(e)$ \ldots absolute time of an event $e$
\end{center}

\noindent
Spanner obeys two rules in order to guarantee this invariant:
\begin{description}
  \item[Start] Coordinator leader for a write $T_i$ assigns a commit timestamp
    $s_i \geq TT.now().latest$, where $TT.now()$ is executed right after the
    arrival event of the commit request for $T_i$.
  \item[Commit Wait] Coordinator leader ensures that clients cannot see data
    committed by $T_i$ until $TT.after\left(s_i\right)$ is true. This ensures
    that the assigned timestamp $s_i$ is less than the absolute commit time of
    $T_i$.~\cite{Corbett:2012}
\end{description}

Figure~\ref{fig:read-write-transaction} illustrates the timeline of a distributed
read-write transaction. There are three participants groups involved, i.e. $P_2$,
$P_3$ and $P_{CL}$ (which is the coordinator leader). In the first step, all three
participants acquire their locks and execute $TT.now()$ right afterwards to obtain
timestamps independently. Every participant logs his timestamp through Paxos and
after this was done, sends its timestamp to the coordinator leader. The
coordinator leader then computes an overall timestamp for the transaction. This
timestamp must be greater than any timestamp the leader has assigned before,
greater or equal to all timestamps received from the participants and greater than
$TT.now().latest$ at the time the coordinator leader received the commit message.
Then, the computed timestamp is logged by the coordinator leader through Paxos
but no participants are notified until the commit wait rule is satisfied. After
commit wait is satisfied, the coordinator leader notifies all other participants
which are allowed to release their locks after receiving this notification.

@@

    \node[right = -0.15 of notify-t2] (release-locks-t2) {};
    \node[right = -0.15 of notify-t3] (release-locks-t3) {};
    \draw[red] (release-locks-t2.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t2.center) -- ++(0, 0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, -0.25);
    \draw[red] (release-locks-t3.center) -- ++(0, 0.25);
  \end{tikzpicture}
  \caption{Timeline of a read-write transaction.}
  \label{fig:read-write-transaction}
\end{figure}

This also explains why, according to Table~\ref{tbl:reads-writes}, Read-write
transactions only involve the \emph{leader} replicas. Only the leader replicas
of each participant communicate with the coordinator leader. Hence, no other
replicas are involved.
